# Regression Trees  {#RT}

This chapter covers supervised machine learning with classification trees.

_______
Introduction to regression trees

<iframe src="https://drive.google.com/file/d/1wo5bBLWNjTV9_viiH1R6tK21wRr65J2T/preview" width="640" height="480"></iframe>
_______

## Classification vs. regression

What is the difference between classification and regression?


* In classification, the response represents a category (e.g. "apples", "oranges", "bananas").

* In regression, the response represents a numeric value (e.g. price of a house).

* **All of the above.**

* None of the above.

______________

## Split the data

These examples will use a subset of the [Student Performance Dataset](https://archive.ics.uci.edu/ml/datasets/Student+Performance) from UCI ML Dataset Repository.

The goal of this exercise is to predict a student's final Mathematics grade based on the following variables: `sex`, `age`, `address`, `studytime` (weekly study time), `schoolsup` (extra educational support), `famsup` (family educational support), `paid` (extra paid classes within the course subject) and `absences`.

The response is `final_grade` (numeric: from 0 to 20, output target).

After initial exploration, split the data into training, validation, and test sets. In this chapter, we will introduce the idea of a validation set, which can be used to select a "best" model from a set of competing models.

In Chapter \@ref(CT), we demonstrated a simple way to split the data into two pieces using the `sample()` function. In this exercise, we will take a slightly different approach to splitting the data that allows us to split the data into more than two parts (here, we want three: train, validation, test). We still use the `sample()` function, but instead of sampling the indices themselves, we will assign each row to either the training, validation or test sets according to a probability distribution.

The dataset `grade` is already in your workspace.

```{r}
url <- "https://assets.datacamp.com/production/repositories/710/datasets/3d720e80d1ad70a88322c2175fa0e6041761a5f9/grade.csv"
if(!file.exists("./Data/grade.csv")){ download.file(url, destfile = "./Data/grade.csv")}
grade <- read.csv("./Data/grade.csv")
```

* Take a look at the data using the `str()` function.

```{r}
# Look/explore the data
str(grade)
```

* Set a seed (for reproducibility) and then sample `n_train` rows to define the set of training set indices.  Draw a sample of size `nrow(grade)` from the number 1 to 3 (with replacement). You want approximately 70% of the sample to be 1 and the remaining 30% to be equally split between 2 and 3.

```{r}
# Randomly assign rows to ids (1/2/3 represents train/valid/test)
# This will generate a vector of ids of length equal to the number of rows
# The train/valid/test split will be approximately 70% / 15% / 15% 
set.seed(1)
assignment <- sample(1:3, size = nrow(grade), prob = c(0.70, 0.15, 0.15), replace = TRUE)
```

* Subset `grade` using the sample you just drew so that indices with the value 1 are in `grade_train`, indices with the value 2 are in `grade_valid`, and indices with 3 are in `grade_test`.

```{r}
# Create a train, validation and tests from the original data frame 
grade_train <- grade[assignment == 1, ]   # subset the grade data frame to training indices only
grade_valid <- grade[assignment == 2, ]   # subset the grade data frame to validation indices only
grade_test <- grade[assignment == 3, ]    # subset the grade data frame to test indices only
```

___________

## Train a regression tree model

In this exercise, we will use the `grade_train` dataset to fit a regression tree using `rpart()` and visualize it using `rpart.plot()`. A regression tree plot looks identical to a classification tree plot, with the exception that there will be numeric values in the leaf nodes instead of predicted classes.

This is very similar to what we did previously in Chapter \@ref(CT). When fitting a classification tree, we use `method = "class"`, however, when fitting a regression tree, we need to set `method = "anova"`. By default, the `rpart()` function will make an intelligent guess as to what the method value should be based on the data type of your response column, but it's recommened that you explictly set the method for reproducibility reasons (since the auto-guesser may change in the future).

The `grade_train` training set is loaded into the workspace.

___________

* Using the `grade_train` dataframe and the given formula, train a regresion tree.

```{r}
# Train the model
grade_model <- rpart(formula = final_grade ~ ., 
                     data = grade_train, 
                     method = "anova")
```

* Look at the model output by printing the model object.

```{r}
# Look at the model output                      
print(grade_model)
```

* Plot the decision tree using `rpart.plot()`.

```{r}
# Plot the tree model
rpart.plot(x = grade_model, yesno = 2)
```

**Each node shows:**

 + the predicted value,
 + the percentage of observations in the node.

_________

Performance metrics for regression

<iframe src="https://drive.google.com/file/d/1T9S-2mEMIzIYbMCC-uOqKLwoKC7SMN9G/preview" width="640" height="480"></iframe>

___________

## Evaluate a regression tree model

Predict the final grade for all students in the test set. The grade is on a 0-20 scale. Evaluate the model based on test set [RMSE (Root Mean Squared Error)](https://en.wikipedia.org/wiki/Root-mean-square_deviation). RMSE tells us approximately how far away our predictions are from the true values.

_________

* First generate predictions on the `grade_test` data frame using the `grade_model` object.

```{r}
# Generate predictions on a test set
pred <- predict(object = grade_model,   # model object 
                newdata = grade_test)   # test dataset
```

* After generating test set predictions, use the `rmse()` function from the `Metrics` package to compute test set RMSE.

```{r}
# Compute the RMSE
rmse(actual = grade_test$final_grade, 
     predicted = pred)
```

_________

What are the hyperparameters for a decision tree?

<iframe src="https://drive.google.com/file/d/1JvNzQHZvDz_ehB6kK1bJGEIufcmxGdGY/preview" width="640" height="480"></iframe>

_________

## Tuning the model

Tune (or "trim") the model using the `prune()` function by finding the best "CP" value (CP stands for "Complexity Parameter").

_______

* Print the CP Table, a matrix of information on the optimal prunings (based on CP).

```{r}
# Plot the "CP Table"
plotcp(grade_model)
# Print the "CP Table"
print(grade_model$cptable)
```

* Retrieve the optimal CP value; the value for CP which minimizes cross-validated error of the model.

```{r}
# Retreive optimal cp value based on cross-validated error
(opt_index <- which.min(grade_model$cptable[, "xerror"]))
(cp_opt <- grade_model$cptable[opt_index, "CP"])
```

* Use the `prune()` function trim the tree, snipping off the least important splits, based on CP.

```{r}
# Prune the model (to optimized cp value)
grade_model_opt <- prune(tree = grade_model, 
                         cp = cp_opt)
                          
# Plot the optimized model
rpart.plot(x = grade_model_opt, yesno = 2)
```

_______

Grid search for model selection

<iframe src="https://drive.google.com/file/d/1c-nCCip2F4ewJ4nqlfDLyTiHFyZ6nxyz/preview" width="640" height="480"></iframe>

_______

## Generate a grid of hyperparameter values

Use `expand.grid()` to generate a grid of `maxdepth` and `minsplit` values.

___________

* Establish a list of possible values for `minsplit` and `maxdepth`.

```{r}
### Establish a list of possible values for minsplit and maxdepth
minsplit <- seq(1, 4, 1)
maxdepth <- seq(1, 6, 1)
```

* Use the `expand.grid()` function to generate a data frame containing all combinations.

```{r}
# Create a data frame containing all combinations 
hyper_grid <- expand.grid(minsplit = minsplit, maxdepth = maxdepth)
```

* Take a look at the resulting grid object.

```{r}
# Check out the grid
head(hyper_grid)

# Print the number of grid combinations
nrow(hyper_grid)
```

_______

## Generate a grid of models

In this exercise, we will write a simple loop to train a "grid" of models and store the models in a list called `grade_models`. R users who are familiar with the `apply` functions in R could think about how this loop could be easily converted into a function applied to a list as an extra-credit thought experiment.

________

* Create an empty list to store the models from the grid search.

```{r}
# Number of potential models in the grid
(num_models <- nrow(hyper_grid))

# Create an empty list to store models
grade_models <- list()
```

* Write a loop that trains a model for each row in `hyper_grid` and adds it to the `grade_models` list.

    + The loop will by indexed by the rows of `hyper_grid`.
    
    + For each row, there is a unique combination of the `minsplit` and `maxdepth` values that will be used to train a model.

```{r}
# Write a loop over the rows of hyper_grid to train the grid of models
for (i in 1:num_models) {

    # Get minsplit, maxdepth values at row i
    minsplit <- hyper_grid$minsplit[i]
    maxdepth <- hyper_grid$maxdepth[i]

    # Train a model and store in the list
    grade_models[[i]] <- rpart(formula = final_grade ~ ., 
                               data = grade_train, 
                               method = "anova",
                               minsplit = minsplit,
                               maxdepth = maxdepth)
}
```

