--- 
title: "[Tree Based Models in R](https://learn.datacamp.com/courses/tree-based-models-in-r)"
author: '[Alan T. Arnholt](https://alanarnholt.github.io/)'
date: 'Last compiled: `r format(Sys.time(), "%b %d, %Y")`'
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Prerequisites

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, fig.align = "center", fig.width = 6, fig.height = 4, message = FALSE, warning = FALSE)
library(tidyverse)
library(caret)
library(mlbench)
library(caTools)
library(openintro)
# Parallel Processing
library(doMC)
registerDoMC(cores = 12)
```

This material is from the [DataCamp](https://www.datacamp.com) course [Tree Based Models in R](https://learn.datacamp.com/courses/tree-based-models-in-r) by Erin LeDell and Gabriela de Queiroz.  Before using this material, the reader should have completed and be comfortable with the material in the DataCamp module [Introduction to R](https://www.datacamp.com/courses/free-introduction-to-r).


Reminder to self: each `*.Rmd` file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'tidyverse'
), 'packages.bib')
```
